Dynamic Modeling, Parameter Estimation and Uncertainty Analysis in \proglang{R}

### 2016-10-03 article.tex
flowchart

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage{amssymb}
\usepackage{enumerate}

@@

For the mathematical description of bile flow, a three-compartment differential equation model is used. TCA is pipetted into the buffer compartment (\code{TCA_buffer}) from where it is transported into the cells (\code{TCA_cell}). Intracellular TCA is exported back to the buffer or the canalicular compartment (\code{TCA_cana}). Finally, canalicular TCA flows back into the buffer compartment. These processes give rise to the following differential equations:
\begin{equation}
	\begin{aligned}
		\frac{\rm d}{ {\rm d}t} \verb!TCA_buffer! &= \verb!-import * TCA_buffer + export_sinus * TCA_cell!\\
		& \quad \verb! + reflux * TCA_cana!\\
		\frac{\rm d}{ {\rm d}t} \verb!TCA_cana! &= \verb!export_cana * TCA_cell - reflux * TCA_cana!\\
		\frac{\rm d}{ {\rm d}t} \verb!TCA_cell! &= \verb!import * TCA_buffer - export_sinus * TCA_cell!\\
		&\quad \verb! - export_cana * TCA_cell!.
	\end{aligned}
	\label{eq:ode}
\end{equation}

@@

The steady-state relation between \code{TCA_cana} and \code{TCA_cell} can be derived analytically from eq.~\eqref{eq:ode}. It reads
\begin{equation}
	\verb!TCA_cana! = \verb!export_cana * TCA_cell / reflux!
	\label{}
\end{equation}
This relation can be explicitly used in a parameter transformation to express \code{TCA_cana} in terms of other parameters. The dimension of the parameter space is thereby reduced by one. The following implementation shows how we would use the existing transformation function \code{p} to generate an alternative transformation function \code{pSS} to replace \code{p} in the prediction function \code{g*x*p}.
\begin{CodeChunk}
\begin{CodeInput}
pSS <- NULL
trafos <- summary(p)
conditions <- names(trafos)
for (n in conditions) {

>>>>

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}
\usepackage{amssymb}
\usepackage{enumerate}
\pdfinclusioncopyfonts=1

@@

For the mathematical description of bile flow, a three-compartment differential equation model is used. TCA is pipetted into the buffer compartment (\code{TCA_buffer}) from where it is transported into the cells (\code{TCA_cell}). Intracellular TCA is exported back to the buffer or the canalicular compartment (\code{TCA_cana}). Finally, canalicular TCA flows back into the buffer compartment. These processes give rise to the differential equations and corresponding flowchart in Figure~\ref{fig:flowchart}.
\begin{figure}[ht]
	\centering
	\begin{minipage}{0.55\textwidth}
		\textbf{A}
		\begin{equation*}
			\begin{aligned}
				\frac{\rm d}{ {\rm d}t} \verb!TCA_buffer! &= \verb!-import * TCA_buffer!\\
				& \quad \verb! + export_sinus * TCA_cell!\\
				& \quad \verb! + reflux * TCA_cana!\\
				\frac{\rm d}{ {\rm d}t} \verb!TCA_cana! &= \verb!export_cana * TCA_cell!\\
				& \quad \verb! - reflux * TCA_cana!\\
				\frac{\rm d}{ {\rm d}t} \verb!TCA_cell! &= \verb!import * TCA_buffer!\\
				& \quad \verb! - export_sinus * TCA_cell!\\
				&\quad \verb! - export_cana * TCA_cell!.
			\end{aligned}
		\end{equation*}
	\end{minipage}
	\begin{minipage}{0.44\textwidth}
		\textbf{B}\\
	\includegraphics[width = \textwidth]{images/flowchart}
	\end{minipage}
	\caption{Differential equations and flowchart of the reaction network. Taurocholic acid,  TCA, is transported between three compartments by four different processes. (A) Assuming mass-action kinetics, the three dynamic states satisfy a set of coupled differential equations. (B) The equations are visualized in a flowchart. }
	\label{fig:flowchart}
\end{figure}

@@

The steady-state relation between \code{TCA_cana} and \code{TCA_cell} can be derived analytically from the ODE, Figure~\ref{fig:flowchart}A. It reads
\begin{equation}
	\verb!TCA_cana! = \verb!export_cana * TCA_cell / reflux!
	\label{}
\end{equation}
This relation can be explicitly used in a parameter transformation to express \code{TCA_cana} in terms of other parameters. The dimension of the parameter space is thereby reduced by one. The following implementation shows how we would use the existing transformation function \code{p} to generate an alternative transformation function \code{pSS} to replace \code{p} in the prediction function \code{g*x*p}.
\begin{CodeChunk}
\begin{CodeInput}
pSS <- NULL
trafos <- summary(p)
conditions <- names(trafos)
for (n in conditions) {
